---
title: "PPPData"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
#install.packages("tidyverse")
#install.packages("tidycensus")

library(tidyverse)
library(lubridate)
library(tidycensus)
```
#
In this R program I will executing queries on the full collection of Paycheck Protection program records (as of 06/30/2021) to analyze these 6 issues:
1) The number and volume of undisbursed loans to individuals applying for loans as single practitioners or sole proprietors.
2) [This is work done outside of the R program] Reviewing online reviews that applicants have posted about "bad actor" lenders identified in (1).
3) [This is work done outside of the R program] Looking into the relationships between agents providing a "fast track" application process (in an app) and the same identified "bad actors" from (1).
4) A review of the predominant business types (based on NAICS classification code) of applicant records whose loans were (as of 06/30/2021) undisbursed.
5) Apparently there are many applicants who asked for loans representing salary protection for the maximum salary rate ($100,000/year). For many applicants reporting their industry-affiliation via the NAICS code, this salary rate appears to be incongruous with the rates reported by the Bureau of Labor Statistics (BLS), and here we identify the application records for single applications requesting the maximum salary rate.
6) Similarly, I am comparing the number of loan applications for individuals in a particular NAICS area (barbers/beauticians/cosmetologists) to the estimated number of existing jobs for those professions provided by BLS to raise a question about the veracity of those applicants reported roles.
7) Back to undisbursed loans - are there particular locations where the applicants' loans wer undisbursed as of 06/30/2021? 

```{r}
mykey <- scan("./censusapikey.txt", what=character())
census_api_key(mykey, install=TRUE, overwrite=TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


Here is where the files are:
https://data.sba.gov/dataset/8aa276e2-6cab-4f86-aca4-a7dde42adf24/resource/be89502c-1961-4a79-a44c-979eb3f411a8/download/public_150k_plus_210630.csv
https://data.sba.gov/dataset/8aa276e2-6cab-4f86-aca4-a7dde42adf24/resource/e796a768-4785-46c1-98a7-8bd3a1ef0ab9/download/public_up_to_150k_1_210630.csv

```{r}
# Read in the data. I prepared the data set from the data downloaded from the SBA
all_loans <- read_rds("data/all_loans.rds")
```

This section is for the analysis

```{r}
# Read in the NAICS codes refrence table
naics_codes <- read_csv('data/naics_codes.csv') %>%
  mutate(NAICSCode= as.character(naics_code)) %>%
  select(NAICSCode, title)

# Then we might want to break that up by NAICS code
all_loans <- all_loans %>% 
  left_join(naics_codes)

# Get a count by Approval amount. We are looking for the sweet spot for maximum amount associated with a single-person business,
# which is around $20,833 (=$100000(per year)/12 (months in a year) * 2.5 months)
# This counts by NAICS code the loans that hover around the single person sweet spot
max1personalLoan <- all_loans %>% 
  filter((CurrentApprovalAmount >=20830) & (CurrentApprovalAmount <=20835)) %>%
  filter(JobsReported == 1) %>%
  group_by(CurrentApprovalAmount, NAICSCode, title) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

# Let's sum up all by NAICS title!
top100NAICS <- max1personalLoan %>%
  group_by(title) %>%
  summarize(total=sum(count)) %>%
  arrange(desc(total)) %>%
  head(100)
  
write_csv(top100NAICS, "data/top100.csv")
```

Here we explore the issues with undisbursed loans to individual applicants (businesses with 1 saved job)
and then undisbursed loans to individual applicants with high salary ($100,000)

```{r}
# Summarize information about undisbursed loans by Originating Lender
all_single_loans_summary <- all_loans %>% 
  filter(JobsReported == 1) %>%
  mutate(UndisbursedAmount = replace_na(UndisbursedAmount, 0)) %>%
  mutate(CurrentApprovalAmount = replace_na(CurrentApprovalAmount, 0)) %>%
  group_by(OriginatingLender) %>%
  summarise(totalUndisbursed = sum(UndisbursedAmount), totalApproved = sum(CurrentApprovalAmount), count = n() ) %>%
  mutate(pctUndisbursed = totalUndisbursed/totalApproved) %>%
  arrange(desc(count))

# Grab the undisbursed loans to applicants listing only a single job reported as being supported
# where the status date is before 45 days from the time the data was published (06/30/2021). This might account
# for loans that were still in process as of the end of June

all_single_undisbursed_loans <- all_loans %>% 
  filter(UndisbursedAmount != 0) %>%
  filter(JobsReported == 1) %>%
  filter(mdy(LoanStatusDate) <= mdy("05/15/2021") )

# Summarize by Originating Lender to identify questionable lenders
all_sdls_summary_by_lender <- all_single_undisbursed_loans %>%
  group_by(OriginatingLender) %>%
  summarise(total = sum(UndisbursedAmount), count = n()) %>%
  arrange(desc(total))

# Summarize by NAICS to review the most frequent NAICS categories
all_sdls_summary_by_naics <- all_single_undisbursed_loans %>%
  group_by(title) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
# Note the predominance of certain types of businesses (barber shops and beauty salons!)
```
This next section calcualtes undisbursed loans by 5-digit ZIP code

```{r}
# Get the 5 digit zip

all_single_undisbursed_loans <- all_single_undisbursed_loans %>%
  mutate(zip5 = str_sub(BorrowerZip, 1, 5))

# Summarize by ZIP code
all_sdls_summary_by_zip5 <- all_single_undisbursed_loans %>%
  group_by(zip5) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# All single disbursed loans

all_single_disbursed_loans <- all_loans %>% 
  filter(UndisbursedAmount == 0) %>%
  filter(JobsReported == 1)

# Summarize by undisbursed amount and by count of loans
all_single_disbursed_summary <- all_single_disbursed_loans %>%
  group_by(OriginatingLender) %>%
  summarise(total = sum(UndisbursedAmount), count = n()) %>%
  arrange(desc(total))

# Undisbursed loans by race and gender
 all_single_undisbursed_loans %>%
  group_by(across(all_of(gbcols))) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
 
 
 # ALl loans to individuals (sole proprietors, individual S Corps/LLCs, etc.)
all_single_loans <- all_loans %>% 
  filter(JobsReported == 1) 

# Group by location - looking for patterns
  
gbcols = c("BorrowerAddress", "BorrowerCity", "BorrowerZip")
all_single_loans %>%
  group_by(across(all_of(gbcols))) %>%
  summarise(total = sum(CurrentApprovalAmount), count = n()) %>%
  arrange(desc(count))
 
# Maybe look at zip codes and get their racial populations??
# Can we get an aggregation of count of loans in Zip Codes where the black population is greater than X%
#
```

THis is legacy code, maybe delete

```{r}
single_beauty_salons <- all_loans %>% 
  filter(NAICSCode == "812112") %>% 
  filter(JobsReported == 1) %>%
  filter((CurrentApprovalAmount >=20830) & (CurrentApprovalAmount <=20835)) 

single_beauty_salons %>%
  filter(UndisbursedAmount != 0) %>%
  group_by(OriginatingLender) %>%
  #summarise(count = n()) %>%
  summarise(total = sum(UndisbursedAmount)) %>%
  arrange(desc(total))

# Single beauty salons that have undisbursed funds with Capital Plus Financial as the Originating Lender
sbs_undisbursed_cpf <- single_beauty_salons %>%
  filter(UndisbursedAmount != 0) %>%
  filter(OriginatingLender == "Capital Plus Financial, LLC")

gbcols <- c("Race", "Gender")

# Undisbursed loans by race and gender
sbs_undisbursed_cpf %>%
  group_by(across(all_of(gbcols))) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Can we group by address?
ppp_by_address <- all_loans %>%
  group_by(BorrowerAddress, BorrowerCity) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

ppp_by_address %>% head(100)

```

